### 2021-06-26 v2-Manuscript/Manuscript.tex
working on a new version with streamlined model

%%%~~~~~~~~~~~~~~~~~ Preamble ~~~~~~~~~~~~~~~~~~~~~~
\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb}

\usepackage{xcolor, hyperref}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\usepackage{geometry}
\geometry{margin=0.65in}
\setlength\parindent{0pt}

\title{Dynamic Knapsack Problem}
\author{AS, WM, KV}
\date{\today}

\allowdisplaybreaks

%%%~~~~~~~~~~~~~~~~~ Document ~~~~~~~~~~~~~~~~~~~~~~
\begin{document}

\maketitle
\tableofcontents
\newpage

\section{Questions/Comments}

\subsection{Questions}
\begin{itemize}
	\item Should $ul$ definition include units violated?? it seems a constraint might be violated otherwise
\end{itemize}

\subsection{Comments}
\begin{itemize}
	\item
\end{itemize}

\section{Assumptions} 
\begin{itemize}
	\item Assume all new patients are arriving at the beginning of the period 
	\begin{itemize}
		\item Assume arrivals follow poisson distribution
	\end{itemize}

	\item It is assumed that patients do not change complexities within some limit ($TL$). 
	\begin{itemize}
		\item Let's assume our limit is 10 periods, patients will never change complexities within the 10 periods.  
		\item However, if a patient waits for 11 periods there is a possibility they increase complexity
	\end{itemize}

	\item We assume patient complexity transitions follow binary distribution - each period a patient can become more complex with a certain probability
	\begin{itemize}
		\item While patient is waiting less that $TL$, the transition probability is 0.
		\item After $TL$, transition probability is some arbitrary value
	\end{itemize}

	\item It is assumed if a patient is scheduled in period one, it means they are served "immediately", regardless if in practice the appointment is at the beginning or end of the period.
	\item We distinguish between two types of reschedules: good and bad reschedules
	\begin{itemize}
		\item Good reschedules are reschedules where a patient is rescheduled to an earlier period
		\item Bad reschedules are reschedules where a patient is rescheduled to a later period
	\end{itemize}

	\item We assume that only specific reschedules are allowed (to simplify the model, and remove  redundancies)
	\begin{itemize}
		\item Good reschedules are allowed from any period after 2 into period 1
		\item Bad reschedules are only allowed from period 1 to any period after 2
	\end{itemize}

	\item It is assumed that there is a certain default expected number of PPE units available for all periods
	\begin{itemize}
		\item However, in the period 1, there is some random deviation from the expected number of units
		\item This random deviation follows some uniform distribution
	\end{itemize}
	\item We allow some violation of PPE units, in rare cases (with high cost), in order to accomodate changes due to variability.
	\begin{itemize}
		\item If we have violation of PPE units - the extra capacity comes from external source
	\end{itemize}
\end{itemize}

\section{MDP Model}
\subsection{Decision Epochs}
Decisions are made at the beginning of each time period. There are 3 relevant time intervals to look at:
\begin{itemize}
	\item Pre-decision state ($S$)
	\begin{itemize}
		\item Pre-decision state defines the initial state on which a decision is required
		\item This information drives decision making
	\end{itemize}

	\item Post-decision state ($\hat{S}$)
	\begin{itemize}
		\item Post-decision state is the time when the decision has been executed, but no new info came in
		\item This state defines the immediate cost of an action
	\end{itemize}

	\item Post-transition state ($S'$)
	\begin{itemize}
		\item This is the state when new information has come in (transition randomness). Defines next pre decision state
		\item It is primarily used to generate expectation for the ADP
	\end{itemize}
\end{itemize}

\subsection{State Space}
\subsubsection{Description}
State is defined by current available and used resources, patient waitlist/demand, and patients already scheduled
\[  \vec{S}  = (\vec{ul}, \vec{pw}, \vec{ps}) \]
\begin{itemize}
	\item $\vec{ul} = ul_{p}$ - Units left over from previous period for resource $p$ plus deviation for that unit for period 1
    \item $\vec{pw} = pw_{mdc}$ - Patients of complexity $d$, CPU $c$, on a wait list for $m$ periods ($m$ of 0 - just arrived)
    \item $\vec{ps} = ps_{tmdc}$ - Patients of complexity $d$, CPU $c$, scheduled to period $t$, waiting for $m$ periods ($m$ of 0 - just arrived)
\end{itemize}

\subsection{Action Sets}
\subsubsection{Description}
Decision consists of rescheduling currently scheduled patients, and scheduling patients on waitlist. There are also some goal and auxiliary variables
\[  \vec{A} = (\vec{sc}, \vec{rsc}, \vec{uv}, \text{auxiliary variables}) \] 
\begin{itemize}
    	\item $\vec{sc} = sc_{tmdc}$ - Patients of complexity $d$, CPU $c$, waiting for $m$ periods, to schedule in period $t$
	\item $\vec{rsc} = rsc_{tt'mdc}$ -Patients of complexity $d$, CPU $c$, waiting for $m$ periods, to reschedule from period $t$ to period $t'$
	\item $\vec{uv} = uv_{tp}$ - goal variable, violation on number of resources used for period $t$, of resource $p$
\end{itemize}

\subsubsection{Auxiliary Variables}
\begin{itemize}
	\item $uvb_{tp}$ - binary variable to enforce $uv$ variable without objective function
	\item $\hat{ul}_{p}$ - post-decision unit leftover at period 1
	\item $ulb_{p}$ - binary variable to enforce $ul$ variable without objective function
	\item $\hat{uu}_{tp}$ - post-decision units used
	\item $\hat{pw}_{mdc}$ - post-decision patients waiting
	\item $\hat{ps}_{tmdc}$ - post-decision patients scheduled
\end{itemize}
\subsubsection{Auxiliary Variable Definition}
\label{auxiliary constraints}
\begin{alignat}{10}
	& \hat{uu}_{tp} 
		&& =  \sum_{mdc} U_{pdc} \hat{ps}_{tmdc} \quad
		&& \forall tp \\ 
	& \hat{pw}_{mdc} 
		&& = pw_{mdc} - \sum_{t} sc_{tmdc} \quad 
		&& \forall mdc \\ 
	& \hat{ps}_{tmdc} 
		&& = ps_{tmdc} + sc_{tmdc} + 
			\sum_{t} rsc_{tt'mdc} - \sum_{t'} rsc_{tt'mdc} \quad 
		&& \forall tmdc
\end{alignat}

\begin{itemize}

	\item Define Resource Violation Variable
	\begin{alignat}{10}
			& \hat{uv}_{tp} 
				&& \le  M(uvb_{tp}) \quad
				&& \forall tp \\
			& \hat{uv}_{1p} 
				&& \le  (\hat{uu}_{1p} - uen_{1p} - ul_{p}) + M(1-uvb_{tp}) \quad
				&& \forall p \\
			& \hat{uv}_{tp} 
				&& \le  (\hat{uu}_{tp} - uen_{tp}) + M(1-uvb_{tp}) \quad
				&& \forall t \in \{ 2..T \} p \\
	\end{alignat}

	\item Define Units Left Over Variable
	\begin{alignat}{10}
			& \hat{ul}_{p} 
				&& \ge 0 \quad 
				&& \forall p \\
			& \hat{ul}_{p} 
				&& \ge uen_{1p} + ul_{p} - \hat{uu}_{1p} \quad 
				&& \forall p \\
			& \hat{ul}_{p} 
				&& \le M (ulb_{p}) \quad 
				&& \forall p \\
			& \hat{ul}_{p} 
				&& \le (uen_{1p} + ul_{p} - \hat{uu}_{1p}) + M(1-ulb_{p}) \quad 
				&& \forall p
	\end{alignat}
\end{itemize}

\subsubsection{State-Action Constraints}
\label{state-action constraints}
\begin{itemize}

	\item Resource Usage Constraint
	\begin{alignat}{10}
			& \hat{uu}_{1p} 
			&& \le uen_{1p} + ul_{p} + uv_{1p} \quad 
			&& \forall p \\
		& \hat{uu}_{tp} 
			&& \le uen_{tp} + uv_{tp} \quad
			&& \forall t \in \{ 2..T \}p 
	\end{alignat}

	\item Custom bounds on when reschedules are allowed
	\begin{alignat}{10}
		& rsc_{tt'mdc} && = 0 \quad && \forall t \in \{ 2...T \}, t' \in \{2...T\} mdc \\
		% & rsc_{tt'mdc} && = 0 \quad && \forall t \in \{ 1 \}, t'\in \{ 3...T \} mdc  \\
		& rsc_{tt'mdc} && = 0 \quad && \forall tt'mdc, \text{where } t=t'=1 
	\end{alignat}

	\item Bounds on Schedules/Reschedules
	\begin{alignat}{10}
		& \sum_{t'} rsc_{tt'mdc} 	&& \le ps_{tmdc} \quad  && \forall tmdc  \\
		& \sum_{t} sc_{tmdc} 		&& \le pw_{mdc} \quad 	&& \forall mdc 
	\end{alignat}

	\item Bounds on states
	\begin{alignat}{10}
		& ul_{p} 	&& \le uen_{p} * 3 \quad 	&& \forall p \\
		& pw_{mdc} 	&& \le pea_{dc} * 20 \quad 			&& \forall mdc \\
		& ps_{tmdc} && \le pea_{dc} * 4 \quad 			&& \forall tmdc
	\end{alignat}
\end{itemize}

\subsection{Transition Probabilities}

\subsubsection{Uncertainty Sources}
\begin{enumerate}

	\item Number of patients arriving this period - $pw_{0dc}$
		\begin{itemize}
			\item $pea_{dc}$ - represents the random number of patients arriving. Follows a poisson distribution.
		\end{itemize}

	\item Transition between patient complexities within the wait list - $pw_{m>=1,dc}$
		\begin{itemize}
			\item $pwt_{mdc}$ represents the random number of patients of complexity $d$, CPU $c$, waiting for $m$ period, who became more complex. Follows binomial distribution.
			\item $ptp_{mdc}$ represents transition probability to a higher complexity
		\end{itemize}

	\item Transition between patient complexities within the scheduled list - $ps_{tmdc}$
		\begin{itemize}
			\item $pst_{tmdc}$ represents the random number of patients of complexity $d$, CPU $c$, waiting for $m$ period, scheduled into $t$ period, who became more complex. Follows binomial distribution.
		\end{itemize}

	\item Amout of expected units of PPE resource for the next time period - $ue_{1p}$
		\begin{itemize}
			\item $ued_{p}$ represents the random deviation of resource $p$ from the expectation. Follows uniform distribution.
			\item $uen_{p}$ expected units of resource $p$ per period
		\end{itemize}
\end{enumerate}

\subsubsection{Transition Constraints}

\begin{enumerate}
	\item Transition from $\vec{ue}$ to $\vec{ue'}$ - Resource Carry Over
	\begin{alignat}{10}
		& ul'_{p} 
			&& = \hat{ul}_{p} + ued_{p} \quad 
			&& \forall p
	\end{alignat}

	\item Transition from $\vec{pw}$ to $\vec{pw'}$ - Flow of patients on waitlist
		\begin{alignat}{10}
			& pw'_{0dc} 
				&& = pea_{dc} \quad
				&& \forall dc \\  
			& pw'_{mdc} 
				&& = \hat{pw}_{m-1,dc} + 
				\overbrace{pwt_{m-1,d-1,c} - pwt_{m-1,dc}}^\text{
					change in complexities} \quad
				&& \forall m \in \{ 1...M-1 \} dc \\  
			& pw'_{Mdc} 
				&& = \sum_{M-1}^{M} \big( 
					\hat{pw}_{mdc} + 
					\overbrace{pwt_{md-1,c} - pwt_{mdc}}^\text{
						change in complexities} \big) \quad
				&& \forall dc
		\end{alignat}

	\item Transition from $\vec{ps}$ to $\vec{ps'}$ - Flow of patiensts scheduled
		\begin{alignat}{10}
			& ps'_{t0dc} 
				&& = 0 \quad
				&& \forall tdc \\  
			& ps'_{Tmdc} 
				&& = 0 \quad
				&& \forall mdc \\  
			& ps'_{tmdc} 
				&& = \hat{ps}_{t+1,m-1,dc} + 
				\overbrace{pst_{t+1,m-1,d-1,c} - pst_{t+1,m-1,dc}}^\text{
					change in complexities} \quad
				&& \forall t \in \{1...T-1\} m \in \{ 1...M-1 \} dc \\  
			& ps'_{tMdc} 
				&& = \sum_{M-1}^{M} \big( 
				\hat{ps}_{t+1mdc} + 
				\overbrace{pst_{t+1,md-1,c} - pst_{t+1,mdc}}^\text{
					change in complexities} \big) \quad
				&& \forall t \in \{1...T-1\} dc
		\end{alignat}
\end{enumerate}

\subsection{Costs}
Cost will come from 4 sources:  
\begin{itemize} 
	\item waiting ($cw$)
	\item rescheduling ($cc$)
	\item scheduling ($cs$)
	\item resource violation ($M$)
\end{itemize}

\begin{equation}\begin{alignedat}{10}
	& C = c(\vec{\hat{pw}}, \vec{\hat{ps}}, \vec{rsc}, \vec{uv}) = 
		&& 
			\overbrace{ \sum_{mdc} cw^{m} (\hat{pw}_{mdc}) }^\text{
				cost of waiting}  + 
			\overbrace{ \sum_{tmdc} cw^{M} (\hat{ps}_{tMdc})}^\text{
				prevent infinite reschedules} + \\
	&	&&	+
				\overbrace{
					\sum_{tmdc} cs^{t} (sc_{tmdc})
				}^\text{Prefer earlier appointments} \\ 
	&	&&  +
			\overbrace{
				(1.5cc) \sum_{
				\substack{tt`mdc \\ t' > t}} 
				rsc_{tt'mdc}}^\text{Bad Reschedule} - 
			\overbrace{ 
				(0.5cc) \sum_{
				\substack{tt`mdc \\ t' < t}}
				rsc_{tt'mdc}}^\text{Good Reschedule} + 
			M \sum_{tp} uv_{tp}
\end{alignedat}\end{equation}

\section{LP Model}

\subsection{Full LP} 
Given a full MDP model, the equivalent LP would look as follows:
\begin{equation}
	\max_{\vec{v}} \sum \alpha (\vec{S}) v(\vec{S}) 
\end{equation}

subject to
\begin{equation}
	c(\vec{\hat{pw}}, \vec{\hat{ps}}, \vec{rsc}, \vec{uv}) + \gamma \sum_{\vec{p}} p(\vec{pea}, \vec{pwt}, \vec{pst}, \vec{ued}) v( \vec{S'} | \vec{S}, \vec{A}, \vec{pea}, \vec{pwt}, \vec{pst}, \vec{ued}) \ge v(\vec{S}) \quad \forall \vec{S} \vec{A}
\end{equation}

\section{Aproximate Dynamic Programming Model}
\subsection{Full ADP Model}
Let's convert it into ADP. We do that by changing $v(\vec{S})$ to an approximation as follows:
\begin{equation}\begin{alignedat}{10}
	& v(\vec{ul}, \vec{pw}, \vec{ps}) 
		&& =  \beta^0 + \sum_{p} \beta_{p}^{ul} ul_{p} +
			\sum_{mdc} \beta_{mdc}^{pw} pw_{mdc}  + 
			\sum_{tmdc} \beta_{tmdc}^{ps} ps_{tmdc}
\end{alignedat} \end{equation}

This gives the following LP
\begin{equation}
	\max_{\vec{\beta}} \sum \alpha (\vec{ul}, \vec{pw}, \vec{ps}) v(\vec{ul}, \vec{pw}, \vec{ps})
\end{equation}
Subject to:
\begin{equation}\begin{alignedat}{10}
\label{full-adp-equation}
	& c(\vec{\hat{pw}}, \vec{\hat{ps}}, \vec{rsc}, \vec{uv}) + \gamma \sum_{\vec{p}} p \Bigg( 
		\beta^0 +
		\bigg( \beta^{ul}_{p} (\hat{ul}_{p} + uen_{d}) \bigg) + \\  
	& 	\bigg( \sum_{dc} \beta_{0dc}^{pw} pea_{dc} \bigg) +
		\bigg( \sum_{m=1,dc}^{M-1} \beta_{mdc}^{pw} 
			(\hat{pw}_{m-1dc} + pwt_{m-1d-1c} - pwt_{m-1dc}) \bigg) + \\
	&	\qquad \qquad \bigg( \sum_{dc} \beta_{Mdc}^{pw} 
			(\sum_{M-1}^{M} \hat{pw}_{mdc} + pwt_{md-1c} - pwt_{mdc})\bigg) + \\
	& 	\bigg( \sum_{tdc} \beta_{t0dc}^{ps} 0 \bigg) +
		\bigg( \sum_{mdc} \beta_{Tmdc}^{ps} 0 \bigg) +
		\bigg( \sum_{t=1m=1dc}^{T-1M-1} \beta_{tmdc}^{ps} 
			(\hat{ps}_{t+1m-1dc} + pst_{t+1m-1d-1c} - pst_{t+1m-1dc}) \bigg) + \\	
	&	\bigg( \sum_{t=1Mdc}^{T-1} \beta_{tmdc}^{ps} 
			(\sum_{M-1}^{M} (\hat{ps}_{t+1mdc} + pst_{t+1md-1c} - pst_{t+1mdc})) \bigg)
	\Bigg) \ge v(\vec{ul}, \vec{pw}, \vec{ps}) \quad \forall \vec{S} \vec{A}
\end{alignedat} \end{equation}
\begin{equation}
	\vec{\beta} \ge 0 \quad \forall \beta
\end{equation}
And all auxiliary constraints in "\nameref{auxiliary constraints}"

\subsection{Expectation ADP Model}
Steps for rearranging and converting the constraint (\ref{full-adp-equation}) into expectation:
\begin{itemize}
	\item $E[ued_p] = 0$
	\item $E[pwt_{mdc}] = ptp_{mdc} * \hat{pw}_{mdc}$ 
	\item $E[pst_{tmdc}] = ptp_{mdc} * \hat{ps}_{tmdc}$ 
\end{itemize}

\begin{equation}
	\beta^{0}: \qquad (1-\gamma) \beta^{0}
\end{equation}
\begin{equation}
\beta^{ul}_{p}: \qquad 
	\sum_{p} \beta_{p}^{ul} \Big( ul_{p} - \gamma ( \hat{ul}_{p}) \Big)
\end{equation}
\begin{equation}\begin{alignedat}{10}
& \beta^{pw}_{mdc}: \qquad 
	&& \sum_{dc} \beta_{0dc}^{pw} \Big( pw_{0dc} - \gamma (pea_{dc}) \Big) + 
	\sum_{m=1dc}^{M-1} \beta_{mdc}^{pw} 
	\Big( pw_{mdc} - \gamma 
		(\hat{pw}_{m-1dc} + E[pwt_{m-1d-1c}] - E[pwt_{m-1dc}]) \Big) + \\
&	&& \sum_{dc} \beta_{Mdc}^{pw} 
	\Big( pw_{Mdc} - \gamma \sum_{M-1}^{M}
		(\hat{pw}_{mdc} + E[pwt_{md-1c}] - E[pwt_{mdc}]) \Big)
\end{alignedat} \end{equation}
\begin{equation}\begin{alignedat}{10}
& \beta^{ps}_{tmdc}: \qquad 
	&& \sum_{tdc} \beta_{t0dc}^{ps} \Big( ps_{t0dc}) \Big) + 
	\sum_{mdc} \beta_{Tmdc}^{ps} \Big( ps_{Tmdc}) \Big) + \\
&	&& 
	\sum_{t=1m=1dc}^{T-1M-1} \beta_{tmdc}^{ps} 
	\Big( ps_{tmdc} - \gamma 
		(\hat{ps}_{t+1m-1dc} + E[pst_{t+1m-1d-1c}] - E[pst_{t+1m-1dc}]) \Big) + \\
&	&& \sum_{t=1dc}^{T-1} \beta_{tMdc}^{ps} 
	\Big( ps_{tMdc} - \gamma \sum_{M-1}^{M}
		(\hat{ps}_{t+1mdc} + E[pst_{t+1md-1c}] - E[pst_{t+1mdc}]) \Big)
\end{alignedat} \end{equation}

Let's say $E[V]$ is the addition of all parts above. Then the ADP model converted to expectation would look as follows:
\begin{equation}
	\max_{\vec{\beta}} \Big( \sum \beta^{0} + \sum \beta^{ut} E[ut] + \sum \beta^{pw} E[pw] + \sum \beta^{ps} E[ps] \Big)
\end{equation}
Subject to:
\begin{equation}
	E[V] \le c(\vec{pw}, \vec{sc}, \vec{rsc}, \vec{uv}) \quad \forall \vec{S} \vec{A} 
\end{equation}

\subsection{ADP Master Problem}
\label{Dual of ADP LP}
Converting to Dual
\begin{equation}
	\min_{w} \sum_{\vec{S} \vec{A}} w(\vec{S} \vec{A}) c(\vec{S} \vec{A})
\end{equation}
Subject To:

\begin{alignat}{10}
	& \beta^{0}: 
		&&	\sum_{\vec{S}\vec{A}}\vec{w} (1 - \gamma) = 1 \\
	& \beta^{ul}: 
		&& 	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ul_{p} - 
			\gamma (\hat{ul}_{p}) \Big) \ge E[ul_{p}]
		&&	\forall p \\
	&	\beta^{pw}: 
		&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(pw_{0dc} - 
		\gamma (pea_{dc}) \Big) \ge E[ps_{0dc}]
		&&	\forall dc \\
	&	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(pw_{mdc} - 
		\gamma (\hat{pw}_{m-1dc} + E[pwt_{m-1d-1c}] - E[pwt_{m-1dc}]) \Big) \ge E[pw_{mdc}]
		&&	\forall m \in \{1...M-1\} dc \\
	&	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(pw_{Mdc} - \gamma 
			\sum_{M-1}^{M} (\hat{pw}_{mdc} + E[pwt_{md-1c}] - E[pwt_{mdc}]) \Big) \ge E[pw_{Mdc}]
		&&	\forall dc \\ 
	&	\beta^{ps}: 
		&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{t0dc} \Big) \ge E[ps_{t0dc}]
		&&	\forall tdc \\
	&	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{Tmdc} \Big) \ge E[ps_{Tmdc}]
		&&	\forall mdc \\
	&	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{tmdc} - \gamma 
			(\hat{ps}_{t+1m-1dc} + E[pst_{t+1m-1d-1c}] - E[pst_{d}]) \Big) \ge E[ps_{tmdc}] \;
		&&	\forall \{1...T-1\} \{1...M-1\}dc \\
	&	&&	\sum_{\vec{S}\vec{A}}\vec{w} \Big(ps_{tMdc} - \gamma \sum_{M-1}^{M}
		(\hat{ps}_{t+1mdc} + E[pst_{t+1md-1c}] - E[pst_{d}]) \Big) \ge E[ps_{tMdc}] \;
		&&	\forall \{1...T-1\} dc
\end{alignat}

\begin{equation}
	\text{State Action Bounds:} \qquad w \ge 0 \qquad \forall w
\end{equation}

\subsection{ADP Pricing Problem}
\label{Pricing Problem}
\begin{equation}
	\min_{(\vec{ue}, \vec{uu}, \vec{pw}, \vec{pe}, \vec{ps}) \in S,  (\vec{sc}, \vec{rsc}) \in A } c(\vec{pw},\vec{sc},\vec{rsc},\vec{uv}) - E[V]
\end{equation}
Subject to:
constraints in sections "\nameref{state-action constraints}" and "\nameref{auxiliary constraints}"

\section{Solution Explanation}
\subsection{Algorithm for solving}
To get $\vec{\beta}$ values, which will be used to generate an action follow steps below:
\begin{enumerate}
	\item Perform a monte-carlo simulation (following some arbitrary policy) to get $E[ue], E[uu], E[pw] E[pw]$
	\item Create an initial feasible set of state-action pairs - $\vec{w}$
	\item Solve model in section "\nameref{Dual of ADP LP}" where each state-action pairs in $\vec{w}$ corresponds to a variable and parameters for all the constraints for a specific action.
	\item Solve model in section "\nameref{Pricing Problem}", where duals from problem in step 3  correspond to $\vec{\beta}$ values. 
		\begin{itemize}
			\item If objective function is less than 0, add solution as a single state-action pair to $\vec{w}$ and go to step 3
			\item If objective function is greater than 0, continue to next step
		\end{itemize}
	\item Duals from problem in step 3 correspond to final $\vec{\beta}$ values
\end{enumerate}

\subsection{Generating an Action}
Once $\vec{\beta}$ values have been approximated, you may use the model below to generate a recommended action for a specific state.
\begin{equation*}
	\min_{\vec{A}} c(\vec{S}, \vec{A}) - \gamma v(S')
\end{equation*}

\begin{equation}\begin{alignedat}{10}
	&	v(S') = \Bigg( 
		&& 	\beta^0 + 
			\bigg( \sum_{p} \beta_{1p}^{ue} 
				(uen_{d} + ue_{1p} - \hat{uu}_{1p} + uv_{1p}) \bigg)+
			\bigg( \sum_{t=2,p}^{T} \beta_{tp}^{ue} uen_{d} \bigg)+ \\
	&	&& 	\bigg( \sum_{tp}^{T-1} \beta_{tp}^{uu} 
				( \hat{uu}_{t+1,p} + \sum_{mdc} E[pst_{tmdc}] (U_{pd+1c} - U_{pdc})) \bigg) + \\
	&	&&	\bigg( \sum_{dc} \beta_{0dc}^{pw} pea_{dc} \bigg) +
			\bigg( \sum_{m=1,dc}^{M-1} \beta_{mdc}^{pw} 
				(\hat{pw}_{m-1dc} + E[pwt_{m-1d-1c}] - E[pwt_{m-1dc}]) \bigg) + \\
	&	&&	\qquad \qquad \bigg( \sum_{dc} \beta_{Mdc}^{pw} 
				(\sum_{M-1}^{M} \hat{pw}_{mdc} + E[pwt_{md-1c}] - E[pwt_{mdc}])\bigg) + \\
	& 	&&	\bigg( \sum_{t=1m=1dc}^{T-1M-1} \beta_{tmdc}^{ps} 
				(\hat{ps}_{t+1m-1dc} + E[pst_{t+1m-1d-1c}] - E[pst_{t+1m-1dc}]) \bigg) + \\	
	&	&&	\qquad \qquad \bigg( \sum_{t=1Mdc}^{T-1} \beta_{tmdc}^{ps} 
				(\sum_{M-1}^{M} (\hat{ps}_{t+1mdc} + E[pst_{t+1md-1c}] - E[pst_{t+1mdc}])) \bigg)
			\Bigg)
\end{alignedat}
\end{equation}

Subject to:
constraints in section "\nameref{state-action constraints}"

\end{document}