### 2021-08-12 v2-Manuscript/sections/data.tex

\section{Data}

\subsection{Sets}
\begin{itemize}
	\item $\vec{m}$ - 
		period defining how long a patient has been waiting so far. 0 represents patients who have arrived. 
		$m \in \{0 ... M\}$
	\item $\vec{t}$ - 
		denotes time of various things - scheduled into, waitlist period, rescheduled from. 
		$t \in \{1 ... T\}$
	\item $\vec{d}$ - 
		denotes complexity of a patient. 
		$d \in \{ 1 ... D \}$
	\item $\vec{k}$ - 
		denotes priority of a patient. 
		$d \in \{ 1 ... K \}$
	\item $\vec{c}$ - 
		denotes surgery type of a patient. 
		$c \in \{ 1 ... C \}$
	\item $\vec{p}$ - 
		denotes resource type. 
		$p \in \{ 1 ... P \}$
\end{itemize}

\subsection{Input Data}
\begin{itemize}
	\item $p\_co_{p}$ - set of resources that carries over between periods
	\item $p\_nco_{p}$ - set of resources that doesn't carry over between periods
	\item $p\_ue_{p}$ - amount of resources expected each period
	\item $p\_ued_{p}$ - Deviation from $p\_ue$. Uniform, defined by lower and upper value. The mean must be zero
	\item $U_{pdc}$ - Usage of resource $p$, of complexity $d$, surgery type $c$
	\item $pea_{dc}$ - Patient Expected Arrival
	\item $TL\_d_{dc}$ - Time Limit after which patient can transition in complexity
	\item $TL\_k_{kc}$ - Time Limit after which patient can transition in priority
	\item $ptp\_c_{dc}$ - Probability to transition to higher complexity
	\item $ptp\_p_{kc}$ - Probability to transition to higher priority
\end{itemize}

\subsection{Model Parameters}
\begin{itemize}
	\item $cw_{d}$ - defines cost of waiting, by priority $d$
	\item $cc$ - additional cost of rescheduling 
	\item $\alpha$ - discount factor to be used in the model
\end{itemize}