### 2021-08-16 v2-Manuscript/sections/mdp_model.tex

\section{MDP Model}

\subsection{Decision Epochs}
Decisions are made at the beginning of each time period. There are 3 relevant time intervals to look at:
\begin{itemize}
	\item Pre-decision state ($S$)
	\begin{itemize}
		\item Pre-decision state defines the initial state on which a decision is required
		\item This information drives decision making
	\end{itemize}

	\item Post-decision state ($\hat{S}$)
	\begin{itemize}
		\item Post-decision state is the time when the decision has been executed, but no new info came in
		\item This state defines the immediate cost of an action
	\end{itemize}

	\item Post-transition state ($S'$)
	\begin{itemize}
		\item This is the state when new information has come in (transition randomness). Defines next pre decision state
		\item It is primarily used to generate expectation for the ADP
	\end{itemize}
\end{itemize}

\subsection{State Space}
\subsubsection{Description}
State is defined by carried over resources, deviation from expected resources,, patient waitlist/demand, and patients already scheduled
\[  \vec{S}  = (\vec{ul}, \vec{pw}, \vec{ps}) \]
\begin{itemize}
	\item $\vec{ul} = ul_{p}$ - Units left over from previous period for resource $p$ plus deviation for that unit for period 1.
	\item $\vec{pw} = pw_{mdkc}$ - Patients of complexity $d$, priority $k$, surgery type $c$, on a wait list for $m$ periods ($m$ of 0 - just arrived)
	\item $\vec{ps} = ps_{tmdkc}$ - Patients of complexity $d$, priority $k$,  surgery type $c$, scheduled to period $t$, waiting for $m$ periods ($m$ of 0 - just arrived)
\end{itemize}

\subsection{Action Sets}
\subsubsection{Description}
Decision consists of rescheduling currently scheduled patients, and scheduling patients on waitlist. There are also some goal and auxiliary variables
\[  \vec{A} = (\vec{sc}, \vec{rsc}, \vec{uv}, \text{auxiliary variables}) \] 
\begin{itemize}
	\item $\vec{sc} = sc_{tmdkc}$ - Patients of complexity $d$, priority $k$,  surgery type $c$, waiting for $m$ periods, to schedule in period $t$
	\item $\vec{rsc} = rsc_{tt'mdkc}$ -Patients of complexity $d$, priority $k$, surgery type $c$, waiting for $m$ periods, to reschedule from $t$ to $t'$
\end{itemize}

\subsubsection{Auxiliary Variables}
\begin{itemize}
	\item $\vec{uv} = uv_{tp}$ - goal variable, violation on number of resources used for period $t$, of resource $p$
	\item $uvb_{tp}$ - binary variable to enforce $uv$ variable without objective function
	\item $\hat{ul}_{p \in p\_co}$ - post-decision unit leftover at period 1
	\item $ulb_{p \in p\_co}$ - binary variable to enforce $ul$ variable without objective function
	\item $\hat{uu}_{tp}$ - post-decision units used
	\item $\hat{pw}_{mdkc}$ - post-decision patients waiting
	\item $\hat{ps}_{tmdkc}$ - post-decision patients scheduled
\end{itemize}
\subsubsection{Auxiliary Variable Definition}
\label{auxiliary constraints}
\begin{alignat}{10}
	& \hat{uu}_{tp} 
	&& =  \sum_{mdkc} U_{pdc} \hat{ps}_{tmdkc} \quad
	&& \forall tp \\ 
	& \hat{pw}_{mdkc} 
	&& = pw_{mdkc} - \sum_{t} sc_{tmdkc} \quad 
	&& \forall mdkc \\ 
	& \hat{ps}_{tmdkc} 
	&& = ps_{tmdkc} + sc_{tmdkc} + 
	\sum_{t} rsc_{tt'mdkc} - \sum_{t'} rsc_{tt'mdkc} \quad 
	&& \forall tmdkc
\end{alignat}

\begin{itemize}

	\item Define Resource Violation Variable
	\begin{alignat}{10}
		& \hat{uv}_{tp} 
		&& \le  M(uvb_{tp}) \quad
		&& \forall tp \\
		& \hat{uv}_{1p} 
		&& \le  (\hat{uu}_{1p} - p\_ue_{1p} - ul_{p}) + M(1-uvb_{tp}) \quad
		&& \forall p \\
		& \hat{uv}_{tp} 
		&& \le  (\hat{uu}_{tp} - p\_ue_{tp}) + M(1-uvb_{tp}) \quad
		&& \forall t \in \{ 2..T \} p \\
	\end{alignat}

	\item Define Units Left Over Variable
	\begin{alignat}{10}
		& \hat{ul}_{p} 
		&& \ge 0 \quad 
		&& \forall p  \in p\_co \\
		& \hat{ul}_{p} 
		&& \ge p\_ue_{1p} + ul_{p} - \hat{uu}_{1p} \quad 
		&& \forall p  \in p\_co\\
		& \hat{ul}_{p} 
		&& \le M (ulb_{p}) \quad 
		&& \forall p  \in p\_co\\
		& \hat{ul}_{p} 
		&& \le (p\_ue_{1p} + ul_{p} - \hat{uu}_{1p}) + M(1-ulb_{p}) \quad 
		&& \forall p  \in p\_co
	\end{alignat}
\end{itemize}

\subsubsection{State-Action Constraints}
\label{state-action constraints}
\begin{itemize}

	\item Resource Usage Constraint
	\begin{alignat}{10}
		& \hat{uu}_{1p} 
			&& \le p\_ue_{1p} + ul_{p} + uv_{1p} \quad 
			&& \forall p \in p\_co \\
		& \hat{uu}_{1p} 
			&& \le p\_ue_{1p} + uv_{1p} \quad 
			&& \forall p \in p\_nco \\
		& \hat{uu}_{tp} 
			&& \le p\_ue_{tp} + uv_{tp} \quad
			&& \forall t \in \{ 2..T \}p 
	\end{alignat}

	\item Custom bounds on when reschedules are allowed
	\begin{alignat}{10}
		& rsc_{tt'mdkc} && = 0 \quad && \forall t \in \{ 2...T \}, t' \in \{2...T\} mdkc \\
		% & rsc_{tt'mdkc} && = 0 \quad && \forall t \in \{ 1 \}, t'\in \{ 3...T \} mdkc  \\
		& rsc_{tt'mdkc} && = 0 \quad && \forall tt'mdkc, \text{where } t=t'=1 
	\end{alignat}

	\item Bounds on Schedules/Reschedules
	\begin{alignat}{10}
		& \sum_{t'} rsc_{tt'mdkc} 	&& \le ps_{tmdkc} \quad  && \forall tmdkc  \\
		& \sum_{t} sc_{tmdkc} 		&& \le pw_{mdkc} \quad 	&& \forall mdkc 
	\end{alignat}

	\item Bounds on states
	\begin{alignat}{10}
		& ul_{p} 	&& \le p\_ue_{p} \quad 			&& \forall p \\
		& pw_{mdkc} 	&& \le pea_{dkc} * 2 \quad 			&& \forall mdkc \\
		& ps_{tmdkc} && \le pea_{dkc} \quad 			&& \forall tmdkc
	\end{alignat}
\end{itemize}

\subsection{Transition}

In this section we will have the following:
\begin{itemize}
	\item $p\_ued$ will represent the random variable from $p\_ued$ distribution 
	\item $pea$ will represent the random variable from $pea$ distribution. 
	\item $pwt$ and $pst$, will represent number transitioned in complexity/priority on waitlist and currently scheduled list.
\end{itemize}

\begin{enumerate}
	\item Transition from $\vec{ul}$ to $\vec{ul'}$ - Resource Carry Over
	\begin{alignat}{10}
		& ul'_{p} 
		&& = \hat{ul}_{p} + p\_ued_{p} \quad 
		&& \forall p \in \{ p\_co \} \\
		& ul'_{p} 
		&& = p\_ued_{p} \quad 
		&& \forall p \in \{ p\_nco \} 
	\end{alignat}

	\item Transition from $\vec{pw}$ to $\vec{pw'}$ - Flow of patients on waitlist
	\begin{alignat}{10}
		& pw'_{0dkc} 
			&& = pea_{dkc} \quad
			&& \forall dc \\  
		& pw'_{mdkc} 
			&& = \hat{pw}_{m-1,dkc}
			&& \forall m \in \{ 1...(TL_{c}-1) \} dkc \\  
		& pw'_{mdkc} 
			&& = \hat{pw}_{m-1,dkc} + 
				\overbrace{pwt\_d_{m-1,d-1,c} - pwt\_d_{d}}^\text{
				change in complexities} +  
				\overbrace{pwt\_k_{k-1} - pwt\_k_{k}}^\text{
				change in priorities}\quad 
			&& \forall m \in \{ TL_{c}...M-1 \} dkc \\  
		& pw'_{Mdc} 
			&& = \sum_{M-1}^{M} \big( 
				\hat{pw}_{mdc} + 
				\overbrace{pwt\_d_{md-1,c} - pwt\_d_{d}}^\text{
				change in complexities} +  
				\overbrace{pwt\_k_{k-1} - pwt\_k_{k}}^\text{
				change in priorities} \big) \quad
			&& \forall dc
	\end{alignat}

	\item Transition from $\vec{ps}$ to $\vec{ps'}$ - Flow of patiensts scheduled
	\begin{alignat}{10}
		& ps'_{t0dkc} 
			&& = 0 \quad
			&& \forall tdkc \\  
		& ps'_{Tmdkc} 
			&& = 0 \quad
			&& \forall mdkc \\  
		& ps'_{tmdkc} 
			&& = \hat{ps}_{t+1,m-1,dkc} \quad 
			&& \forall t \ne T m \in \{ 1... TL-1 \} dkc \\ 
		& ps'_{tmdkc} 
			&& = \hat{ps}_{t+1,m-1,dkc} + 
			\overbrace{pst\_d_{t+1,m-1,d-1,c} - pst\_d_{d}}^\text{
				change in complexities}  + 
			\overbrace{pst\_k_{k-1} - pst\_k_{k}}^\text{
				change in priorities} \quad 
			&& \forall t \ne T m \in \{ TL_{c}...M-1 \} dkc \\  
		& ps'_{tMdkc} 
			&& = \sum_{M-1}^{M} \big( 
			\hat{ps}_{t+1mdc} + 
			\overbrace{pst\_d_{t+1,md-1,c} - pst\_d_{d}}^\text{
				change in complexities} + 
			\overbrace{pst\_k_{k-1} - pst\_k_{k}}^\text{
				change in priorities} \big) \quad
			&& \forall t \ne T dkc
	\end{alignat}
\end{enumerate}

\subsection{Costs}

\begin{itemize}
	\item $cs_{tk} = \sum_t (\alpha^t * cw_{k})$
\end{itemize}

\begin{equation}\begin{alignedat}{10}
		& C = c(\vec{\hat{pw}}, \vec{\hat{ps}}, \vec{rsc}, \vec{uv}) = 
		&& 
		\overbrace{ \sum_{mdc} cw_{k} (\hat{pw}_{mdkc}) }^\text{
			cost of waiting}  + 
		\overbrace{
			\sum_{tmdc} cs_{tk} (sc_{tmdkc})
		}^\text{Prefer earlier appointments} \\ 
		&	&&  +
		\overbrace{
			\sum_{
				\substack{tt`mdc \\ t' > t}} 
			(cs_{t'-t, k} + cc) * (rsc_{tt'mdc})}^\text{Bad Reschedule} - 
		\overbrace{ 
			\sum_{
				\substack{tt`mdc \\ t' < t}}
			(cs_{t-t', k} - cc) * (rsc_{tt'mdc})}^\text{Good Reschedule} + 
		M \sum_{tp} uv_{tp}
\end{alignedat}\end{equation}